---
title: "WA 2015 Drug Overdose Hospitalization"
author: "Mamadou Ndiaye, Epidemiologist, WA DOH"
output:
  html_document: default
  html_notebook: default
css: tufte2.css
---


```{r, echo=F,message=FALSE, warning=FALSE}  


library(tidyverse)
library(haven)
library(forcats)
library(caret)
library(Hmisc)
library(purrr)
library(knitr)
library(pander)
library(injuryepi)

load("Y:/Confidential/HSQA/CHS/IVP-Epi/CHARS/.R_Cache/chars_pub2015diag_q4@.RData")
load("Y:/Confidential/HSQA/CHS/IVP-Epi/CHARS/.R_Cache/chars_pub2015diag_f@.RData")




# SOAR::Attach()

options(scipen = 12)
knitr::opts_chunk$set(
  echo    = TRUE,
  comment = NA,
  message = FALSE,
  error   = FALSE)

```  

## Finding All Drug Overdose  


Based on the ICD-9 CM and ICD-10 CM (fourth quarter) in the principal diagnosis 


The results by quarters:

```{r}

pander(filter(chars_pub2015diag_f, drugs == 1) %>% group_by(quarters) %>% count,justify = "left")

```

## Finding All Opioid Overdose



### Opioid Principal Diagnosis Only  


Count of opioid as main diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_pub2015diag_f, opioid == 1) %>% group_by(quarters) %>% count,justify = "left")

```

### Opioid As any Diagnosis   


Count of opioid as any diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_pub2015diag_f, opioid_m == 1) %>% group_by(quarters) %>% count,justify = "left")

```

## Finding Non- Heroin Opioid Overdose


### Non- Heroin Principal Diagnosis Only  


Count of Non-Heroin opioid as main diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_pub2015diag_f, opioid_nh == 1) %>% group_by(quarters) %>% count,justify = "left")

```

### Non-Heroin Opioid As any Diagnosis   


Count of non-heroin opioid as any diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_pub2015diag_f, opioid_nh_m == 1) %>% group_by(quarters) %>% count,justify = "left")

```

## Finding All Heroin Overdose


### heroin Principal Diagnosis Only  

Count of heroin as main diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_pub2015diag_f, heroin == 1) %>% group_by(quarters) %>% count,justify = "left")

```

### heroin As any Diagnosis   

Count of heroin as any diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_pub2015diag_f, heroin_m == 1) %>% group_by(quarters) %>% count,justify = "left")

```
