---
title: "WA 2011-2016 Drug Overdose Hospitalization"
author: "Mamadou Ndiaye, Epidemiologist, WA DOH"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 3
    toc_float: true
css: tufte2.css
---


```{r, echo=F,message=FALSE, warning=FALSE}  


library(tidyverse)
library(haven)
library(forcats)
library(caret)
library(Hmisc)
library(purrr)
library(knitr)
library(pander)
library(injuryepi)
library(highcharter)
library(forecast)
#library(ggplot2)

load("Y:/Confidential/HSQA/CHS/IVP-Epi/CHARS/.R_Cache/chars_pub2015diag_q4@.RData")
load("Y:/Confidential/HSQA/CHS/IVP-Epi/CHARS/.R_Cache/chars_pub2015diag_f@.RData")
load("Y:/Confidential/HSQA/CHS/IVP-Epi/CHARS/.R_Cache/chars_2016@.RData")
load("Y:/Confidential/HSQA/CHS/IVP-Epi/CHARS/.R_Cache/mat@.RData")
load("Y:/Confidential/HSQA/CHS/IVP-Epi/CHARS/.R_Cache/mat_op@.RData")
load("Y:/Confidential/HSQA/CHS/IVP-Epi/CHARS/.R_Cache/mat_nhop@.RData")
load("Y:/Confidential/HSQA/CHS/IVP-Epi/CHARS/.R_Cache/mat_hero@.RData")



# SOAR::Attach()

options(scipen = 12)
knitr::opts_chunk$set(
  echo    = TRUE,
  comment = NA,
  message = FALSE,
  error   = FALSE,
  warning = FALSE)

```  

  The introduction of ICD-10 CM in October 2015 and the use of CDC proposed definitions may have artificially lead to an apparent steeper decrease of drug overdose. However a time series decomposition from 2011 to 2016 showed seasonal effects that manifested in a lower incidence of drug overdose in the first and fourth quaters over the years.  
  In addition this document contrast the results between using the principal diagnosis only and any diagnosis field when capturing the opioid overdose codes for 2016. 

# Finding All Drug Overdose  


Based on the ICD-10 CM **T36-T50** in the principal diagnosis 

```{r eval=F, echo=F}
library(tidyverse)
source("https://raw.githubusercontent.com/epinotes/InjuryEpi/master/add_drug_10cm")
source("https://raw.githubusercontent.com/epinotes/InjuryEpi/master/create_cond_diag")
source("https://raw.githubusercontent.com/epinotes/InjuryEpi/master/create_diag")

```


```{r, eval = F}

library(injuryepi)
chars_2016 <- chars_2016 %>% add_drug_10cm(., uid = "seq_no", main = "diag1")

```
 __The results by quarters__:

```{r, echo=F}

pander(filter(chars_2016, drugs == 1) %>% group_by(quarters) %>% count,justify = "left")

```  


# Finding All Opioid Overdose

Based on `.opioid_10cm = "^T40[01234]X|^T406[09]"`

## Opioid Principal Diagnosis Only  


```{r, eval=F}
chars_2016 <- chars_2016 %>% mutate(opioid = create_cond_diag(.,expr = .opioid_10cm, colvec = 79, cond.var = drugs))

```
Count of opioid as main diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_2016, opioid == 1) %>% group_by(quarters) %>% count,justify = "left")

```

## Opioid As any Diagnosis   

```{r, eval=F}
d_m <- grep("^diag[0-9]|^ecode[0-9]", names(chars_2016), value = F) 

chars_2016 <- chars_2016 %>% mutate(opioid_m = create_cond_diag(.,expr = .opioid_10cm, colvec = d_m, cond.var = drugs))

```
Count of opioid as any diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_2016, opioid_m == 1) %>% group_by(quarters) %>% count,justify = "left")

```

# Finding Non- Heroin Opioid Overdose

Based on `.opioid_nh_10cm = "^T40[0234]X|^T406[09]"`

## Non- Heroin Principal Diagnosis Only  


```{r, eval=F}
chars_2016 <- chars_2016 %>% mutate(opioid_nh = create_cond_diag(.,expr = .opioid_nh_10cm, colvec = 79, cond.var = drugs))

```
Count of Non-Heroin opioid as main diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_2016, opioid_nh == 1) %>% group_by(quarters) %>% count,justify = "left")

```

## Non-Heroin Opioid As any Diagnosis   

```{r, eval=F}
d_m <- grep("^diag[0-9]|^ecode[0-9]", names(chars_2016), value = F) 

chars_2016 <- chars_2016 %>% mutate(opioid_nh_m = create_cond_diag(.,expr = .opioid_nh_10cm, colvec = d_m, cond.var = drugs))

```
Count of non-heroin opioid as any diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_2016, opioid_nh_m == 1) %>% group_by(quarters) %>% count,justify = "left")

```

# Finding All Heroin Overdose

Based on `.heroin_10cm = "^T401X"`

## heroin Principal Diagnosis Only  


```{r, eval=F}
chars_2016 <- chars_2016 %>% mutate(heroin = create_cond_diag(.,expr = .heroin_10cm, colvec = 79, cond.var = drugs))

```
Count of heroin as main diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_2016, heroin == 1) %>% group_by(quarters) %>% count,justify = "left")

```

## heroin As any Diagnosis   

```{r, eval=F}
d_m <- grep("^diag[0-9]|^ecode[0-9]", names(chars_2016), value = F) 

chars_2016 <- chars_2016 %>% mutate(heroin_m = create_cond_diag(.,expr = .heroin_10cm, colvec = d_m, cond.var = drugs))

```
Count of heroin as any diagnosis by quarters:

```{r, echo=F}
pander(filter(chars_2016, heroin_m == 1) %>% group_by(quarters) %>% count,justify = "left")

```

# Monthly Time Series Decomposition from 2011 to 2016  

The decomposition of the time series was done with the R function `stl()`(Seasonal Decomposition of Time Series by Loess). It displays the _original series_, and its decomposition into _seasonal_, _trend(smoothed with loess)_ and _remainder_. 

The graphs were produced with _R packages_ `ggplot2` and `highcharter`

## Time series for all drugs  
  

### __with ggplot2__:  


```{r}
ggplot2::autoplot(stl(mat, s.window = "periodic", robust = TRUE))

```  

### __With highcharter__ (allowing interactivity):  

```{r}  
highcharter::hchart(stl(mat, s.window = "periodic", robust = TRUE))

```



## Time series for all opioids  
 
### __with ggplot2__:  


```{r echo=F}
ggplot2::autoplot(stl(mat_op, s.window = "periodic", robust = TRUE))

```  

### __With highcharter__ : 


```{r echo=F}  
highcharter::hchart(stl(mat_op, s.window = "periodic", robust = TRUE))

```  

## Time series for non-heroin opioid  

### __with ggplot2__:  


```{r echo=F}
ggplot2::autoplot(stl(mat_nhop, s.window = "periodic", robust = TRUE))

```  
### __With highcharter__ : 

  

```{r echo=F} 

highcharter::hchart(stl(mat_nhop, s.window = "periodic", robust = TRUE))

```

## Time series for heroin 

### __with ggplot2__:  


```{r echo=F}
ggplot2::autoplot(stl(mat_hero, s.window = "periodic", robust = TRUE))

```  

### __With highcharter__ : 

 
  

```{r echo=F} 

highcharter::hchart(stl(mat_hero, s.window = "periodic", robust = TRUE))

```
